HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-18 paper/main.tex
New abstract trying to get best of both previous abstracts

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
%\newcommand{\XXX}[1]{}
\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
%\newcommand{\oa}[1]{}
\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
%\newcommand{\bh}[1]{}

@@

Human evaluation of machine translation is a cognitively difficult task and when
annotators rate or rank complete sentences, we see low levels of agreement especially
for longer sentences.  
%In order to extract more reliable annotations, previous research broke down evaluation into smaller components by asking humans to annotate semantic roles on the source and translation, and then to align them. 
In order to extract more reliable annotations, previous research broke down
evaluation into smaller components in various ways. 
We propose a new evaluation method, HUME, 
which is based on the Universal Conceptual Cognitive Annotation (UCCA)
semantic annotation.
UCCA is applicable across many languages, it has better coverage of important
semantic phenomena than previously used semantic annotation,
semantic role labeling, and it is grounded in the words in the sentence. 
We apply UCCA only over the source or reference sentence, and not over the
possibly garbled translation.
%, 
%and this makes annotation more reliable and also efficient as we can reuse labor
%intensive semantic annotations.
We show that HUME can be reliably applied to
four different target languages and that sentence-level inter-annotator agreement
does not decrease over longer sentences.

@@

Human judgement is the central, and perhaps most fundamental, criterion for
estimating the quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and fluency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence. This
is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
%as a function of
with
the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated, and hence cannot inform developers in overcoming these errors.

These problems are partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}). For automatic measures,
these are commonly words or n-grams, for manual measures some structural
information is taken into account \parcite{machacek:bojar:segranks:2015}.
% the following is specific to automatic measures:
%quantifying the overlap of its sub-parts relative a reference translation.
A promising line of research decomposes
%More recent work on MT evaluation proposed measures
%that decompose
over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is
HMEANT \cite{lo2011structured}.
\XXX{Do we want to cut the following shorter?}
These measures can localize semantic errors in the translation, but
their focus on verb argument structures misses many pervasive semantic phenomena
(such as inter-clause linkage and nominal argument structures) and thus they only provide
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

@@

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
Importantly, HUME only requires semantically annotating the source sentence,
thus avoiding the annotation of machine-generated text 
and allowing re-use of the source semantic annotation for measuring the quality
of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
%An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.

After a brief review (\secref{sec:background}), we describe HUME in detail
(\secref{sec:hume}) and highlight its benefits over HMEANT
(\secref{sec:hmeant_comp}).
Our experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech (\secref{sec:experiments}) document the
efficiency (time of annotation) and inter-annotator agreement scores for HUME.

% We conduct experiments with four language pairs: English-German, English-Polish,
% English-Romanian and English-Czech, and report efficiency (time of annotation) and
% inter-annotator agreement scores (\secref{sec:experiments}).
% 
% \XXX{The following should go rather to the conclusion}
% Efficiency results that HUME
% can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,
% which is only about five times slower than MEANT, and hence still practical to
% employ on a moderate scale. HUME's agreement results are on par with those
% of standard sentence-level ranking measures, but unlike them,
% HUME's scores do not decrease with longer sentences.\bh{We may
%   have to focus more on qualitative aspects, if we just try to sell
%   the measure on IAA, then the figures are not stunning. OA: is that OK?}

@@

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is generally done by ranking the outputs of multiple systems
e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
individual adequacy/fluency scores to each translation, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation
is not a scalable solution.
%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a manually created reference translation.
This decomposition further allows system developers
to localize the errors, or at least unconfirmed parts within the translation.
The most commonly used measures decompose over n-grams or individual words, e.g., 
BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR  \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments,
their decomposition over words and n-grams is better suited for localizing errors
in the output of shallow MT methods, and is less instructive in identifying the semantic
nature of the errors.

>>>>

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
%\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
\newcommand{\XXX}[1]{}
%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\oa}[1]{}
%\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
\newcommand{\bh}[1]{}

@@

Human evaluation of machine translation normally uses sentence-level measures such as relative ranking
or adequacy/fluency scales. However these provide no insight into the MT errors,
 and do not scale well with sentence 
length. We argue for a semantics-based evaluation, which aims to capture how much of the meaning
is retained in the MT output, and can support the construction and tuning of semantics-based MT. 
We present a novel human semantic evaluation measure for MT, Human
UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA) semantic representation scheme.
Making progress over previous work, the proposed measure takes into account
a wider range of semantic phenomena and does not rely on semantic annotation
of the MT output itself, whose interpretation is often unclear.
We experiment with four language pairs, demonstrating HUME's broad applicability,
as well as its reliability, reflected in its inter-annotator agreement rates.

%Human evaluation of machine translation is a cognitively difficult task and when
%annotators rate or rank complete sentences, we see low levels of agreement especially
%for longer sentences.  
%In order to extract more reliable annotations, previous research broke down evaluation into smaller 
%components by asking humans to annotate semantic roles on the source and translation, and then to align them. 
%In order to extract more reliable annotations, previous research broke down
%evaluation into smaller components in various ways. 
%We propose a new evaluation method
%HUME
%which is based on semantic annotation called Universal Conceptual Cognitive Annotation (UCCA). UCCA
%is applicable across many languages, it provides a complete structure over sentences, and it is 
%grounded in the words in the sentence. 
%We apply UCCA only over the source or reference sentence, and not over the
%possibly garbled translation.
%, 
%and this makes annotation more reliable and also efficient as we can reuse labor
%intensive semantic annotations.
%We show that HUME can be reliably applied to
%four different target languages and that sentence level inter-annotator agreement
%does not decrease over longer sentences.

@@

Human judgement is the central, and perhaps most fundamental, criterion for
estimating the quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and fluency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence. This
is indeed reflected in the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated, and hence cannot inform developers in overcoming these errors.

These problems were partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}),
most commonly over its words or n-grams,
quantifying the overlap of its sub-parts relative a reference translation.
More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is
HMEANT \cite{lo2011structured}.
These measures can localize semantic errors in the translation, but
their focus on verb argument structures misses many pervasive semantic phenomena
(such as inter-clause linkage and nominal argument structures) and thus they only provide
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

@@

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
Importantly, HUME only requires semantically annotating the source sentence,
thus avoiding the annotation of machine-generated text 
and allowing re-use of the source semantic annotation for measuring the quality
of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
An elaborate comparision of HUME and HMEANT is given in \secref{sec:hmeant_comp}.

We conduct experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores (\secref{sec:experiments}). Efficiency results that HUME
can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,
which is only about five times slower than MEANT, and hence still practical to
employ on a moderate scale. HUME's agreement results are on par with those
of standard sentence-level ranking measures, but unlike them,
HUME's scores do not decrease with longer sentences.\bh{We may
  have to focus more on qualitative aspects, if we just try to sell
  the measure on IAA, then the figures are not stunning. OA: is that OK?}

@@

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is generally done by ranking the outputs of multiple systems
e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
individual adequacy/fluency scores to each translation, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Most commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a manually created reference translation.
This decomposition further allows system developers
to localize the errors, or at least unconfirmed parts within the translation.
The most commonly used measures decompose over n-grams or individual words, e.g., 
BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR  \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments,
their decomposition over words and n-grams is better suited for localizing errors
in the output of shallow MT methods, and is less instructive in identifying the semantic
nature of the errors.
